---
title: "Functionality of the OpenImageR package"
author: "Lampros Mouselimis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functionality of the OpenImageR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


The following notes and examples illustrate the image processing functions of the OpenImageR package.


## Basic functions

#### readImage

The *readImage* function reads images from a variety of types such as 'png', 'jpeg', 'jpg' or 'tiff',

```{r, eval = T}

library(OpenImageR)

path = file.path(getwd(), 'vignette_1', 'image1.jpeg')

im = readImage(path)
dim(im)

```
<br>


#### imageShow

The *imageShow* function utilizes either a shiny application (if the image is a character path) or the grid.raster function of the base grid package (if the image is a 2- or 3-dimensional object) to display images,
<br>
<br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

imageShow(im)

```
<br><br>


#### writeImage

The *writeImage* function writes a 2- or 3-dimensional object (matrix, data frame or array) in a user specified image format. The supported types are .png, .jpeg, .jpg, .tiff.
<br>
<br>
```{r, eval = F}

writeImage(im, file_name = 'my_image.jpeg')

```
<br>


#### rgb_2gray

To convert an image from RGB to gray one can use the *rgb_2gray* function, which takes a single argument (matrix, data frame or array),
<br>
<br>
```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

path = file.path(getwd(), 'vignette_1', 'image2.jpg')

im = readImage(path)

imageShow(im)
```
<br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}
r2g = rgb_2gray(im)

imageShow(r2g)

```
<br><br>


#### cropImage

The *cropImage* function reduces the size of the image horizontally and vertically. The function takes four arguments : *image* (2- or 3-dimensional object), *new_width* (the desired new width), *new_height* (the desired new height) and *type*. While the type  *'equal_spaced'* crops the image equally in both directions (horizontal, vertical) towards the center of the image, the type *'user_defined'* allows the user to specify which regions of the image should be kept
<br>
<br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

eq_sp = cropImage(im, new_width = 200, new_height = 200, type = 'equal_spaced')

imageShow(eq_sp)

```
<br>


```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

us_def = cropImage(im, new_width = 20:225, new_height = 5:185, type = 'user_defined')

imageShow(us_def)

```
<br><br>

#### flipImage

A [flipped image](https://en.wikipedia.org/wiki/Flipped_image) (or reversed image) is a static or moving image that is generated by a mirror-reversal of an original across a horizontal axis (a flopped image is mirrored across the vertical axis). The first image shows the original image and the second one a horizontal flip,
<br>
<br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

path = file.path(getwd(), 'vignette_1', 'image1.jpeg')

im = readImage(path)

imageShow(im)

```
<br><br>


```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

flp_vert = flipImage(im, mode = 'horizontal')

imageShow(flp_vert)

```
<br><br>

#### rotateFixed

The *rotateFixed* function rotates an image by specific angles (90, 180, 270). I added this function (besides the *rotateImage* function) in the package, because it takes less time to return the output,
<br>
<br>


```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

r270 = rotateFixed(im, 270)

imageShow(r270)

```
<br><br>


#### rotateImage

The *rotateImage* function is more flexible in comparison to the *rotateFixed* function. It allows the user to rotate an image by a specific angle (between 0 and 360 degrees) using either the 'nearest' or the 'bilinear' interpolation method,
<br><br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}


r45 = rotateImage(im, 45, threads = 1)

imageShow(r45)

```
<br><br>

#### resizeImage

The *resizeImage* function takes advantage of two different interpolation methods ( nearest neighbors, bilinear ) to either down- or upsample an image,
<br><br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

path = file.path(getwd(), 'vignette_1', 'image2.jpg')

im = readImage(path)

intBl = resizeImage(im, width = 100, height = 100, method = 'bilinear', normalize_pixels = TRUE)

imageShow(intBl)

```
<br><br>

It is known that resizing an image using bilinear interpolation gives better results, however the nearest neighbors method is faster.
<br>

A further option to resize an image is by using *gaussian_blur* ( down-sampling gives better results than up-sampling that's why is the only option in the OpenImageR package ),
<br>


```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

intGbl = down_sample_image(im, factor = 2.5, gaussian_blur = T)

imageShow(intGbl)

```
<br>

The factor of the function applies equally to the horizontal and vertical dimensions of the image.
<br><br>



#### translation

**translation** is the shifting of an object's location by adding/subtracting a value to/from the X or Y coordinates,

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}


tr = translation(im, shift_rows = 50, shift_cols = -50)

imageShow(tr)

```
<br>


## Image filtering

#### edge_detection

According to Wikipedia, edge detection is the name for a set of mathematical methods which aim at identifying points in a digital image at which the image brightness changes sharply or, more formally, has discontinuities. The *edge_detection* function uses one of the *Frei_chen*, *LoG (Laplacian of Gaussian)*, *Prewitt*, *Roberts_cross*, *Scharr* or *Sobel* filters to perform edge detection,
<br>


```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

path = file.path(getwd(), 'vignette_1', 'image1.jpeg')

im = readImage(path)

edsc = edge_detection(im, method = 'Scharr', conv_mode = 'same')

imageShow(edsc)

```
<br>



#### uniform_filter

In a *uniform filter* all the values within the filter (kernel) have the same weight,


```{r echo = T, eval = T}

kernel_size =  c(4,4)

unf = uniform_filter(im, size = kernel_size, conv_mode = 'same')

unif_filt = matrix(1, ncol = kernel_size[1], nrow = kernel_size[2])/(kernel_size[1] * kernel_size[2])
unif_filt

```
<br><br>



#### image_thresholding 

According to Wikipedia, *thresholding* is the simplest method of image segmentation. From a grayscale image, thresholding can be used to create binary images,


```{r echo = F, eval = F}

# for some reason the following vignette example throws an error during checking. I tried all different ways I even added a similar test case 
# (see the 1st. test in the 'test-gamma_threshold.R' file). Set "eval = F" for this specific vignette example.

```


```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = F}

path = file.path(getwd(), 'vignette_1', 'image2.jpg')

im = readImage(path)

thr = image_thresholding(im, thresh = 0.5)           # if the input image is 3-dimensional it will be converted internally to a matrix

imageShow(thr)

```
<br><br>


#### gamma_correction

*Gamma correction*, or often simply gamma, is the name of a nonlinear operation used to encode and decode luminance or tristimulus values in video or still image systems (Wikipedia),


```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

gcor = gamma_correction(im, gamma = 2)        # show image with gamma correction 

imageShow(gcor)

```
<br><br>


#### ZCAwhiten

[**Whitening** (or sphering)](http://ufldl.stanford.edu/wiki/index.php/Whitening) is the preprocessing needed for some algorithms. If we are training on images, the raw input is redundant, since adjacent pixel values are highly correlated. The purpose of whitening is to reduce the correlation between features and to return features with the same variance,

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

res = ZCAwhiten(im, k = 20, epsilon = 0.1)

imageShow(res)

```
<br><br>


#### delationErosion

[**Dilation** and **erosion**](https://www.mathworks.com/help/images/morphological-dilation-and-erosion.html) are the most basic morphological operations. Dilation adds pixels to the boundaries of objects in an image, while erosion removes pixels on object boundaries,

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

res_delate = delationErosion(im, Filter = c(8,8), method = 'delation')

imageShow(res_delate)

```
<br><br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

res_erosion = delationErosion(im, Filter = c(8,8), method = 'erosion')

imageShow(res_erosion)

```
<br><br>


#### Augmentation

Augmentations are specific transformations applied to an image. The **Augmentation** function allows the user to *flip*, *crop*, *resize*, *shift*, *rotate*, *zca-whiten* and *threshold* an image. It either returns a single image,  

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

path = file.path(getwd(), 'vignette_1', 'image1.jpeg')

im = readImage(path)

dim(im)

augm = Augmentation(im, flip_mode = 'horizontal', crop_width = 20:460, crop_height = 30:450, 
                    
                    resiz_width = 180, resiz_height = 180, resiz_method = 'bilinear', 
                    
                    shift_rows = 0, shift_cols = 0, rotate_angle = 350, 
                    
                    rotate_method = 'bilinear', zca_comps = 100, 
                    
                    zca_epsilon = 0.1, image_thresh = 0.0, verbose = T)

imageShow(augm)

```
<br><br>

or multiple transformed images using either pre-specified or random parameters (here I utilized the lapply function),

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

# random rotations
samp_rot = sample(c(seq(5, 90, 30), seq(270, 350, 30)), 3, replace = F)

# random shift of rows
samp_shif_rows = sample(seq(-50, 50, 10), 3, replace = F)

# random shift of columns
samp_shif_cols = sample(seq(-50, 50, 10), 3, replace = F)



res = lapply(1:length(samp_rot), function(x) 
  
  Augmentation(im, flip_mode = 'horizontal', crop_width = 20:460, crop_height = 30:450, 
               
               resiz_width = 180, resiz_height = 180, resiz_method = 'bilinear', 
                    
               shift_rows = samp_shif_rows[x], shift_cols = samp_shif_cols[x], 
               
               rotate_angle = samp_rot[x], rotate_method = 'bilinear', zca_comps = 100, 
              
               zca_epsilon = 0.1, image_thresh = 0.0, verbose = F))


print(length(res))

imageShow(res[[1]])

```
<br><br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

imageShow(res[[2]])

```
<br><br>

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

imageShow(res[[3]])

```
<br><br>


## Image recognition

#### HOG ( histogram of oriented gradients )

The histogram of oriented gradients (HOG) is a feature descriptor used in computer vision and image processing for the purpose of object detection. The technique counts occurrences of gradient orientation in localized portions of an image. This method is similar to that of edge orientation histograms, scale-invariant feature transform descriptors, and shape contexts, but differs in that it is computed on a dense grid of uniformly spaced cells and uses overlapping local contrast normalization for improved accuracy (Wikipedia).
<br>

The *HOG* function of the OpenImageR package is a modification and extention of the [findHOGFeatures](https://github.com/lastlegion/SimpleCV/blob/develop/SimpleCV/ImageClass.py) function ( SimpleCV computer vision platform ), please, consult the COPYRIGHT file.
<br><br>

The purpose of the function is to create a vector of HOG descriptors, which can be used in classification tasks. It takes either RGB (they will be converted to gray) or gray images as input,
```{r, eval = T}

path = file.path(getwd(), 'vignette_1', 'image2.jpg')

image = readImage(path)

image = image * 255

hog = HOG(image, cells = 3, orientations = 6)
hog

```
<br>

#### HOG_apply

The *HOG_apply* function uses the previous mentioned HOG function to return the HOG-descriptors for the following objects :

* a **matrix** of images such as the mnist data set, where each row represents a different digit (28 x 28 image)
* an **array**, where each slice represents a different image
* a **folder of images** where each file is a different image

In the following code chunk I'll apply the HOG function to an array of images. The result is a matrix, where each row respresents the HOG-descriptors for each array slice (image),

```{r, eval = T}

path_im1 = file.path(getwd(), 'vignette_1', 'image1.jpeg')
path_im2 = file.path(getwd(), 'vignette_1', 'image2.jpg')

tmp_im1 = readImage(path_im1)
tmp_im2 = readImage(path_im2)

tmp_im1 = resizeImage(tmp_im1, 200, 200, normalize_pixels = TRUE)
tmp_im2 = resizeImage(tmp_im2, 200, 200, normalize_pixels = TRUE)

tmp_gray1 = rgb_2gray(tmp_im1)
tmp_gray2 = rgb_2gray(tmp_im2)
dim(tmp_gray2)

tmp_arr = array(0, c(nrow(tmp_gray1), ncol(tmp_gray1), 2))
tmp_arr[,,1] = tmp_gray1
tmp_arr[,,2] = tmp_gray2

res = HOG_apply(tmp_arr, cells = 2, orientations = 3)
res

```


### [image hashing](http://www.hackerfactor.com/blog/?/archives/432-Looks-Like-It.html)

The image hashing functions (*average_hash*, *dhash*, *phash*, *invariant_hash*, *hash_apply*) of the OpenImageR package are implemented in the way **perceptual hashing** works. Perceptual hashing is the use of an algorithm that produces a fingerprint of images (in OpenImageR those fingerprints are *binary features* or *hexadecimal hashes*). The difference between cryptographic and image hashing is that the latter tries to find similar and not exact matches. In cryptographic hashing small differences of the hashes lead to entirely different output, which is not the case for perceptual hashing. A practical application of image hashing would be to compare a database of already created image hashes with a new hash (image) to find similar images in the database.

The *average_hash*, *dhash* and *phash* functions of the OpenImageR package are modifications and extentions of the [ImageHash](https://github.com/JohannesBuchner/imagehash) python library, please, consult the COPYRIGHT file.
<br><br>


###### average_hash

The **average hash** algorithm of the OpenImageR package works in the following way : 1st we convert to grayscale, 2nd we reduce the size of an image (for instance to an 8x8 image), 3rd we average the resulting colors (for an 8x8 image we average 64 colors), 4th we compute the bits by comparing if each color value is above or below the mean and 5th we construct the hash. The result of the *average_hash* function can be either a **hexadecimal hash** (string) or **binary features**,

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

path = file.path(getwd(), 'vignette_1', 'view1.jpg')

image = readImage(path)

imageShow(image)

image = rgb_2gray(image)

aveg_hash = average_hash(image, hash_size = 8, MODE = 'hash', resize = "bilinear")
aveg_hash

aveg_bin = average_hash(image, hash_size = 8, MODE = 'binary', resize = "bilinear")
as.vector(aveg_bin)

```
<br>


###### phash

While the average_hash is fast, it can generate false-misses if there is a gamma correction or a color histogram is applied to the image. The **phash** algorithm extends the average_hash by using the *discrete cosine transform* to reduce the frequencies, 

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

path = file.path(getwd(), 'vignette_1', 'view2.jpg')

image2 = readImage(path)

imageShow(image2)

image2 = rgb_2gray(image2)

ph_hash = phash(image2, hash_size = 8, highfreq_factor = 4, MODE = 'hash', resize = "bilinear")
ph_hash

ph_bin = phash(image2, hash_size = 8, highfreq_factor = 4, MODE = 'binary', resize = "bilinear")
as.vector(ph_bin)

```
<br>


###### [dhash](http://www.hackerfactor.com/blog/index.php?/archives/529-Kind-of-Like-That.html)


In comparison to average_hash and phash, the **dhash** algorithm takes into consideration the difference between adjacent pixels. In the same way as with the average_hash, the resulting hash won't change if the image is scaled or the aspect ratio changes. Increasing or decreasing the brightness or contrast, or even altering the colors won't dramatically change the hash value. Even complex adjustments like gamma corrections and color profiles won't impact the result,

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

path = file.path(getwd(), 'vignette_1', 'view3.jpg')

image3 = readImage(path)

imageShow(image3)


image3a = rgb_2gray(image3)

dh_hash = dhash(image3a, hash_size = 8, MODE = 'hash', resize = "bilinear")
dh_hash

dh_bin = dhash(image3a, hash_size = 8, MODE = 'binary', resize = "bilinear")
as.vector(dh_bin)
```
<br>

By adding a gamma correction we can check if the hash value or the binary features of the *view3.jpg* will be altered,
```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

tmp_image3 = gamma_correction(image3, gamma = 0.5)

imageShow(tmp_image3)

tmp_image3 = rgb_2gray(tmp_image3)

dh_hash_a = dhash(tmp_image3, hash_size = 8, MODE = 'hash', resize = "bilinear")
dh_hash_a

dh_bin_a = dhash(tmp_image3, hash_size = 8, MODE = 'binary', resize = "bilinear")
as.vector(dh_bin_a)

```
<br>


###### invariant_hash

**invariant_hash** is an extension function for image hashing. It takes two images as input (image1, image2) and by altering one of those (random flipping, rotating and cropping) it calculates the *hamming distance* (if the mode is 'binary') or the *levenshtein distance* if (the mode is 'hash'). If any of the *flip*, *rotate*, *crop* equals TRUE then the function returns the MIN,MAX similarity between the two images. If, on the other hand all *flip*, *rotate*, *crop* equal FALSE then a single similarity value is returned meaning no random transformations of the second image are performed.
<br><br>

Although, in the previous example the gamma correction doesn't influence the dhash of the *view3.jpg*, a *horizontal flip* of the image would change both the hash value and the binary features considerably,

```{r fig.width = 1.0, fig.height = 1.0, echo = T, eval = T}

image3b = flipImage(image3, mode = "horizontal")

imageShow(image3b)

image3b = rgb_2gray(image3b)

dh_hash_b = dhash(image3b, hash_size = 8, MODE = 'hash', resize = "bilinear")
dh_hash_b

dh_bin_b = dhash(image3b, hash_size = 8, MODE = 'binary', resize = "bilinear")
as.vector(dh_bin_b)

```
<br>

By using the *invariant hash* in this case with all the available transformations (at the cost of computational time) we can obtain minimum and maximum values for the hamming or the levenshtein distance, 

```{r, eval = T}


inv_hash = invariant_hash(image3a, image3b, mode = 'binary', flip = T, rotate = T, 
                          
                          angle_bidirectional = 10, crop = T)
inv_hash

inv_bin = invariant_hash(image3a, image3b, mode = 'hash', flip = T, rotate = T, 
                         
                         angle_bidirectional = 10, crop = T)
inv_bin

```


In both cases (hash, binary) a minimum value of 0 indicates perfect matches between the two images among the transfomations.


###### hash_apply

The **hash_apply** function applies the single *average_hash*, *phash*, *dhash* to either a matrix, array or folder of images. This function is practical in case of a matrix of images such as the mnist, or the cifar_10 data sets, where each line of the matrix corresponds to an image. Furthermore, if a folder includes many images, then the *hexadecimal hash mode* should be prefered, as it doesn't require as much storage space as the binary mode. The following example illustrates how *hash values* and *binary features* can be computed from a folder of images,


```{r, eval = T}

path = paste0(getwd(), '/TEST_hash/')


hapl_hash = hash_apply(path, hash_size = 6, method = "dhash", mode = "hash", threads = 1,  resize = "nearest")
hapl_hash                # returns both the names of the images and the hash values

hapl_bin = hash_apply(path, hash_size = 6, method = "dhash", mode = "binary", threads = 1,  resize = "nearest")
hapl_bin$files             # names of the images
dim(hapl_bin$hash)         # dimensions of the resulted matrix
head(hapl_bin$hash)        # binary features

```
<br>


```{r, eval = T, echo = F}

# remove cache once vignettes are built

# unlink("The_OpenImageR_package_files", recursive = TRUE)

```
